\GVN{}-Hoist: Hoisting Computations from Branches

### 2016-11-01 hoist.tex
move files around

\maketitle
\begin{abstract}
  Code-hoisting identifies identical computations across the program and hoists
  them to a common dominator so as to save code size.  Not only code-hoisting
  saves code size, a) It exposes redundancies in some cases, b) It exposes more
  instruction level parallelism in the basic block a computation is hoisted, and
  c) It enables other passes like LICM to remove more redundancies.

@@

Compiler techniques to remove redundant computations are composed of an analysis
phase that detects identical computations in the program and a transformation
phase that reduces the number of run-time computations.  Classical scalar
optimizations like CSE \cite{dragonbook} work very well on single basic blocks but
when it comes to detect redundancies across basic blocks these techniques fall
short: more complex passes like GCSE and \PRE{} have been designed to handle these
cases based on dataflow analysis \cite{morel1979global}.  At first these
techniques were described in the classical data-flow analysis framework, and
later the use of the \SSA{} representation lowered the cost in terms of compilation
time \cite{briggs1994effective,chow1997new,kennedy1999partial} and brought these
techniques in the main stream: nowadays \SSA{} based \PRE{} is available in every
industrial compiler.

@@

\begin{itemize}
\item to reduce the code size of the program;
\item to improve function inlining heuristics: functions become cheaper to
  inline when thier code size decreases because inliner heuristics in \LLVM{}
  depends on instruction count.
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, instruction schedulers can move heavy
  computations earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting expressions out of branches, code hoisting can effectively reduce the
  amount of code to be speculatively executed and can reduce the critical path;
\item to reduce register pressure: by moving computations closer to the
  definitions of their operands;
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted and sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

>>>>

\maketitle
\begin{abstract}
  Code-hoisting identifies identical computations across the program and hoists
  them to a common dominator so as to save code size.  Not only code-hoisting
  saves code size, it exposes redundancies in some cases, it exposes more
  instruction level parallelism in the basic block a computation is hoisted, and
  it enables other passes like LICM to remove more redundancies.

@@

Compiler techniques to remove redundant computations are composed of an analysis
phase that detects identical computations in the program and a transformation
phase that reduces the number of run-time computations.  Classical scalar
optimizations like CSE \cite{dragonbook} work very well on single basic blocks and
when it comes to detect redundancies across basic blocks these techniques fall
short: more complex passes like GCSE and \PRE{} have been designed to handle these
cases based on dataflow analysis \cite{morel1979global}.  At first these
techniques were described in the classical data-flow analysis framework, and
later the use of the \SSA{} representation lowered the cost in terms of compilation
time \cite{briggs1994effective,chow1997new,kennedy1999partial} and brought these
techniques in the main stream: nowadays \SSA{} based \PRE{} is available in every
industrial compiler.

@@

\begin{itemize}
\item to reduce the code size of the program;
\item to improve function inlining heuristics: functions become cheaper to
  inline when their code size decreases because inliner heuristics in \LLVM{}
  depends on instruction count.
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, instruction schedulers can move heavy
  computations earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting expressions out of branches, code hoisting can effectively reduce the
  amount of code to be speculatively executed and can reduce the critical path;
\item to reduce register pressure: by moving computations closer to the
  definitions of their operands;
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted and sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

### 2016-11-01 other/libcxx-slides.tex
move files around

>>>>

  \item clang+libc++ vs. gcc+libstdc++ performance update. Constructor of vector in libstdc++ contains memcpy
    in the final assembly.

Benchmark	                        Time (ns)	CPU (ns)	Iterations	Time (ns)	CPU (ns)	Iterations	%(Time)
BM_ConstructIterIter/vector_char/1024	2125	2127	283762	67	67	10205902	3071.641791
BM_ConstructIterIter/vector_size_t/1024	382	383	1830386	223	223	3138652	71.30044843
BM_ConstructIterIter/vector_string/1024	557109	557641	1235	354504	354834	1980	57.15168235

### 2016-11-01 slides/gvn-hoist.tex
move files around

\frame{\frametitle{\GVN{}-Hoist: Hoisting Computations from Branches}
  \begin{itemize}
    \item Identifies identical computations in a function
    \item Hoist identical computations to a common dominator
    \item Uses GVN infrastructure of LLVM
    \item Reduces code size
  \end{itemize}
}

\frame{\frametitle{Open source tools}
  \begin{itemize}
    \item Compiler: gcc, llvm
    \item Gprof
    \item Linux Perf
    \item Create\_gcov
    \item Gooda
  \end{itemize}
}

\frame{\frametitle{PGO: Compiler instrumentation}
  \begin{itemize}
    \item Compiler inserts probes in the code.
    \item Instrumented program collects profile as it executes.
    \item Gives very precise profile information.
    \item Good for benchmarking, analyzing small programs.
    \item gprof/gcov used to process sampled output and display.
  \end{itemize}
}

\frame{\frametitle{PGO: Compiler instrumentation (Useful Commands)}
  gprof:
  \begin{itemize}
    \item gcc -fprofile-generate test.cpp
    \item gprof [exe] [gmon.out] $>$ [outfile]
  \end{itemize}
  gcov:
  \begin{itemize}
    \item gcc -fprofile-arcs -ftest-coverage test.cpp
    \item gcov [options] test.cpp
  \end{itemize}
}

\frame{\frametitle{AutoFDO: Non Intrusive}
  \begin{itemize}
    \item Linux perf collects profiles (Sample based)
    \item No compiler instrumentation, debug-info required while running.
    \item Negligible overhead (1-2\%) [Google]
    \item Run on different machines and at different time intervals. 
    \item Good for system wide profiling.
  \end{itemize}
}

\frame{\frametitle{AutoFDO: Non Intrusive (Useful Commands)}
  \begin{itemize}
    \item perf record/report/stat
    \item collecting profile: perf record -b [exe] [-o perf.data]
    \item report: perf report [perf.data]
  \end{itemize}
}

\frame{\frametitle{Results [Baptiste Wicht et al.]}
    \begin{tabular}{| l | p {2.5cm} | p {3cm} | }
        \hline
        PGO Kind & Avg. Run Time Overhead (\%) & Performance Improvement (\%) \\ \hline
        Instrumented & 16 (highest 53) & 7 to 14  \\ \hline
        AutoFDO & 1 to 2 & 5 to 10 \\ \hline
        AutoFDO+LTO & 1 to 2 & 10 to 30 \\ \hline
    \end{tabular}
}

\frame{\frametitle{Challenges for non-intrusive AutoFDO}
  \begin{itemize}
    \item Value based profiling not available.
    \item Only works on recent Intel machines.
    \item Tradeoff between accuracy of profile and overhead.
    \item Goal is to be as close to the instrumented profile.
  \end{itemize}
}

\frame{\frametitle{Challenges for non-intrusive AutoFDO...}
  \begin{itemize}
    \item Calculating missing edge/basic-block frequencies based on Equivalence class.
    \item Calculating accurate profile of execution paths sharing single line of code based on DWARF discriminator [Baptiste Wicht et al.].
  \end{itemize}
}

\frame{\frametitle{create\_gcov}
    \begin{itemize}
        \item Creates coverage file from perf.data to be used by the compiler.
        \item optimize with profile information: gcc -fauto-profile=file.gcov -O2 test.cpp
        \item Source: https://github.com/google/autofdo
     \end{itemize}
}

\frame{\frametitle{gooda}
    \begin{itemize}
        \item gooda-analyzer
        \item gooda-visualizer
        \item Source: https://github.com/David-Levinthal/gooda
        \item Requires patching linux perf to link against libpfm4 [See: gooda-analyzer/README] 
        \item Requires LBR (Available in Intel Sandy Bridge, Ivy Bridge, Westmere)
    \end{itemize}
}

\frame{\frametitle{Interesting Areas}
  Tools related:
  \begin{itemize}
    \item Extend PGO infrastructure in gcc/llvm:
      currently PGO benefits inliner, code layout, register allocation.
    \item Compiler cost model based on profile to enable demand driven optimization.
    \item Online optimizations driven by hardware performance monitoring [Schneider et al.].
  \end{itemize}
  Architecture specific:
  \begin{itemize}
    \item No AutoFDO tool for generating value profiles.
    \item No LBR for ARM.
  \end{itemize}

}

\frame{\frametitle{References}
    \begin{itemize}
    \item \url{http://static.googleusercontent.com/media/research.google.com/en//pubs/archive/36575.pdf}
    \item Baptiste Wicht et al. ``Hardware Counted Profile-Guided Optimization.''
    \item \url{http://www.burningcutlery.com/derek/docs/instant-profiling-CGO13.pdf}
    \item \url{https://github.com/David-Levinthal/gooda}
    \item \url{https://github.com/google/autofdo}
    \end{itemize}
}

>>>>

\frame{\frametitle{\GVN{}-Hoist: Hoisting Computations from Branches}
  \begin{itemize}
    \item Identifies identical computations in a function
    \item Hoist identical computations to a common dominator
    \item Uses GVN and MemorySSA
    \item Reduces code size
  \end{itemize}
}

\frame{\frametitle{References}
    \begin{itemize}
    \end{itemize}
}
